Untargeted control doc

### 2023-10-05 untargeted/paruntargeted_control.Rmd
Added some basic BMIS visuals

iso_areas <- iso_mzs %>%
  split(.$filename) %>% 
  imap(function(file_peaks, file_name_i){
    msdata <- open_dataset("tmzMLs/pqds") %>%
      filter(amendment==!!amendment) %>%
      filter(polarity==!!polarity) %>%
      filter(filename==file_name_i) %>%
      dplyr::collect()
    file_peaks %>%
      pmap(function(...){
        row_data <- data.frame(...)
        area <- msdata[mz%between%pmppm(row_data$labeled_mzmed, 10)] %>%
          .[rt%between%c(row_data$rtmin/60, row_data$rtmax/60)] %>%
          .[,sum(int)]
        cbind(row_data, area=area)
      }) %>%
      bind_rows()
  }, .progress=TRUE) %>%
  bind_rows() %>%
  mutate(labeled_n_count=n_count-1) %>%
  select(compound_name, filename, area, labeled_n_count)

output_csv_write(iso_areas)
```

```{r individual isotope EIC}
single_cmpd_iso_req <- iso_mzs %>%
  filter(compound_name=="Taurine") %>%
  distinct(compound_name, n_count, labeled_mzmed) %>%
  mutate(labeled_mzmed=round(labeled_mzmed, 5)) %>%
  summarise(req_cmd=paste0("mz%between%pmppm(", labeled_mzmed, ", 10)", collapse="|")) %>%
  pull(req_cmd)
full_arrow_req <- paste(
  'open_dataset("tmzMLs/pqds") %>%',
  'filter(amendment==!!amendment) %>%',
  'filter(samp_type=="Smp") %>%',
  'filter(polarity==!!polarity) %>% ',
  'filter(', single_cmpd_iso_req, ') %>% ',
  'dplyr::collect()'
)
single_cmpd_iso_eic <- eval(parse(text=full_arrow_req))

>>>>

iso_areas <- iso_mzs %>%
  split(.$filename) %>% 
  imap(function(file_peaks, file_name_i){
    msdata <- open_dataset("tmzMLs/pqds") %>%
      filter(amendment==!!amendment) %>%
      filter(polarity==!!polarity) %>%
      filter(filename==file_name_i) %>%
      dplyr::collect()
    file_peaks %>%
      pmap(function(...){
        row_data <- data.frame(...)
        area <- msdata[mz%between%pmppm(row_data$labeled_mzmed, 10)] %>%
          .[rt%between%c(row_data$rtmin/60, row_data$rtmax/60)] %>%
          .[,sum(int)]
        cbind(row_data, area=area)
      }) %>%
      bind_rows()
  }, .progress=TRUE) %>%
  bind_rows() %>%
  mutate(labeled_n_count=n_count-1) %>%
  select(compound_name, filename, area, labeled_n_count)
output_csv_write(iso_areas)
```

# BMIS and quant

```{r get IS areas from Skyline}
n_only_is <- c("L-Isoleucine, 15N", "AMP, 15N5", "GMP, 15N5", 
               "Nitrate, 15N", "DL-Histidine, 15N")
if(system("untargeted/IS_integrations/SkylineRunner.exe", show.output.on.console = FALSE)){
  # Website = https://skyline.ms/wiki/home/software/Skyline/page.view?name=SkylineInstall_64_21-1&submit=false
  stop("Please grab a copy of SkylineRunner from the website")
}
if(!file.exists("untargeted/IS_integrations/GMP_pos.csv")){
  success_vec <- "untargeted/IS_integrations" %>%
  list.files(pattern = "sky$", full.names = TRUE) %>%
  sapply(skyline_files, function(skyfile){
    polarity <- str_extract(skyfile, "POSITIVE|NEGATIVE")
    polarity <- switch(polarity, "POSITIVE"="pos", "NEGATIVE"="neg")
    amendment <- str_extract(skyfile, "Amm|Nit|GMP|Arg")

    sky_cmd <- paste0(
      "untargeted/IS_integrations/SkylineRunner.exe",
      " --in=", skyfile,
      ' --report-name="Molecule Peak Boundaries Modified"', 
      ' --report-file=untargeted/IS_integrations/', amendment, "_", polarity, ".csv"
    )
    failed <- system(sky_cmd)
    if(failed){
      warning(paste("Extraction failed for file", skyfile))
    }

    new_csv_name <- paste0("untargeted/IS_integrations/", amendment, "_", polarity, ".csv")
    new_csv <- read.csv(new_csv_name)
    new_csv$polarity <- polarity
    new_csv$amendment <- amendment
    write.csv(new_csv, new_csv_name, row.names = FALSE)
  })
}
IS_areas_all <- "untargeted/IS_integrations/" %>%
  list.files(pattern = "(Amm|Nit|Arg|GMP)_(pos|neg).csv", full.names = TRUE) %>%
  lapply(read.csv) %>%
  do.call(what = "rbind") %>%
  mutate(Area=as.numeric(Area)) %>%
  mutate(Area=ifelse(is.na(Area), 0, Area)) %>%
  select(filename=File.Name, IS_name=Molecule, rtmin=Min.Start.Time,
         rtmax=Max.End.Time, IS_area=Area, polarity, amendment) %>%
  mutate(filename=str_replace(filename, ".raw", ".mzML")) %>%
  filter(!IS_name%in%n_only_is) %>%
  filter(!IS_name=="Succinic acid, 2H4") %>%
  mutate(lead_IS_area=lead(IS_area), lag_IS_area=lag(IS_area)) %>%
  left_join(parametadata_all) %>%
  mutate(IS_area=case_when(
    shortname=="Amm-Surf-Eve-T3-A"~(lead_IS_area+lag_IS_area)/2,
    shortname=="Nit-Deep-Eve-T0-A"~(lead_IS_area+lag_IS_area)/2,
    TRUE~IS_area
  )) %>%
  select(IS_name, filename, IS_area)

# write.csv(IS_areas_all, "untargeted/IS_integrations/IS_areas.csv", row.names = FALSE)

IS_areas_all %>%
  slice(1) %>%
  left_join(parametadata_all)
  filter(amendment==!!amendment) %>%
  arrange(timestamp) %>%
  mutate(shortname=fct_inorder(shortname)) %>%
  ggplot() +
  geom_col(aes(x=shortname, y=IS_area, fill=samp_type), width=1) +
  facet_wrap(~IS_name, scales="free_y", ncol=3)
```

```{r get PheTau areas from raw files}

```

# Isotope visualization

```{r load isotope mzs and areas}
all_stans <- read_csv("metadata/all_stans.csv") %>%
  filter(polarity==!!polarity) %>%
  filter(compound_type!="Internal Standard") %>%
  select(compound_name, formula)
manual_anno <- readxl::read_excel("untargeted/manual_annotations.xlsx") %>%
  filter(polarity==!!polarity) %>%
  filter(amendment==!!amendment) %>%
  select(feature, compound_name)
best_feat_data <- output_folder %>%
  paste0("best_feat_data.csv") %>%
  read_csv() %>%
  left_join(manual_anno) %>%
  left_join(all_stans)
iso_mzs <- read_csv(paste0(output_folder, "iso_mzs.csv"))
iso_areas <- read_csv(paste0(output_folder, "iso_areas.csv"))
```

```{r individual isotope EIC}
single_cmpd_iso_req <- iso_mzs %>%
  filter(compound_name=="Guanine") %>%
  mutate(labeled_mzmed=round(labeled_mzmed, 5)) %>%
  summarise(req_cmd=paste0("mz%between%pmppm(", labeled_mzmed, ", 10)", collapse="|")) %>%
  pull(req_cmd)
full_arrow_req <- paste(
  'open_dataset("tmzMLs/pqds") %>%',
  'filter(amendment==!!amendment) %>%',
  'filter(samp_type=="Smp") %>%',
  'filter(polarity==!!polarity) %>% ',
  'filter(', single_cmpd_iso_req, ') %>% ',
  'dplyr::collect()'
)
single_cmpd_iso_eic <- eval(parse(text=full_arrow_req))
